State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-05-27 StateOfPractice/Papers/PainPoints_OS_MI/PainPoints_OS_MI.tex
Split versions of SOP papers

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with Harvard style bibliographic references

%\documentclass[preprint,12pt,authoryear]{elsarticle}
%\documentclass[3p, 12pt,authoryear]{elsarticle}
\documentclass[final, 3p, times, authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath,amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

@@

%\usepackage{tikz}
%\usetikzlibrary{mindmap}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{pbox}

%% Comments
\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}} %Spencer Smith
\newcommand{\jc}[1]{\authornote{red}{JC}{#1}} %Jacques Carette
\newcommand{\mn}[1]{\authornote{magenta}{MN}{#1}} %Mike Noseworthy
\newcommand{\ad}[1]{\authornote{cyan}{AD}{#1}} %Ao Dong

\newcommand{\notdone}[1]{\textcolor{red}{#1}}
\newcommand{\done}[1]{\textcolor{black}{#1}}

@@

\journal{Academic Radiology}

@@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={}, 
%%            city={},
%%            postcode={}, 
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{State of the Practice for Medical Imaging Software}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[CAS]{Spencer Smith}
\author[CAS]{Ao Dong}
\author[CAS]{Jacques Carette}
\author[ECE]{Mike Noseworthy}

\affiliation[CAS]{organization={McMaster University, Computing and Software Department},%Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\affiliation[ECE]{organization={McMaster University, Electrical Engineering},%Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

@@

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
%\begin{highlights}
%\item Research highlight 1
%\item Research highlight 2
%\end{highlights}

\begin{keyword}
	medical imaging, research software, software engineering, software
	quality, Analytic Hierarchy Process, developer interviews
\end{keyword}

\end{frontmatter}

%% \linenumbers

@@

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[scale=0.25]{figures/MPR.png}        
    \end{center}
    \caption{Example brain image showing a multi-planar reformat using Horos
	(free open-source medical imaging/DICOM viewer for OSX, based on OsiriX)}
    \label{Fig_Example}
\end{figure}

@@

We designed a general methodology to assess the state of the practice for
research software \citep{SmithEtAl2021, SmithAndMichalski2022}. Details can be
found in Section~\ref{ch_methods}.  Our methodology has been applied to MI
software \citep{Dong2021} and Lattice Boltzmann Solvers \citep{SmithEtAl2022,
Michalski2021}.  This methodology builds off prior work to assess the state of
the practice for such domains as Geographic Information Systems
\citep{smith2018state}, Mesh Generators \citep{smith2016state}, Seismology
software \citep{Smith2018Seismology}, and Statistical software for psychology
\citep{smith2018statistical}.  In keeping with the previous methodology, we have
maintained the constraint that the work load for measuring a given domain should
be feasible for a team as small as one person, and for a short time, ideally
around a person month of effort. We consider a person month as $20$ working days
($4$ weeks in a month, with $5$ days of work per week) at $8$ person hours per
day, or $20 \times 8 = 160$ person hours.

@@

\begin{figure}[!ht]
\includegraphics[scale=0.66]{figures/template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure}

@@

\section{Measurement Results} \label{ch_results}

Table~\ref{tab_final_list} shows the 29 software packages that we measured,
along with summary data collected in the year 2020. We arrange the items in
descending order of LOC. We found the initial release dates (Rlsd) for most
projects and marked the two unknown dates with ``?''. The date of the last
update is the date of the latest update, at the time of measurement. We found
funding information (Fnd) for only eight projects.  For the Number Of
Contributors (NOC) we considered anyone who made at least one accepted commit as
a contributor. The NOC is not usually the same as the number of long-term
project members, since many projects received change requests and code from the
community.  With respect to the OS, 25 packages work on all three OSs: Windows
(W), macOS (M), and Linux (L). Although the usual approach to cross-platform
compatibility was to work natively on multiple OSes, five projects achieved
platform-independence via web applications. The full measurement data for all
packages is available in \citet{Dong2021-Data}.

\begin{table}[!ht]
\centering
\begin{tabular}{p{6cm}lllllllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Rlsd} & \multirow{2}{*}{Updated} & \multirow{2}{*}{Fnd} & \multirow{2}{*}{NOC} & \multirow{2}{*}{LOC} & \multicolumn{3}{c}{OS} & \multirow{2}{*}{Web} \\ \cline{7-9}
 &  &  &  &  &  & W & M & L &  \\ \midrule
ParaView \citep{Ahrens2005} & 2002 & 2020-10 & \checkmark & 100 & 886326 & \checkmark & \checkmark & \checkmark & \checkmark \\
Gwyddion \citep{Nevcas2012} & 2004 & 2020-11 &  & 38 & 643427 & \checkmark & \checkmark & \checkmark &  \\
Horos \citep{horosproject2020} & ? & 2020-04 &  & 21 & 561617 &  & \checkmark &  &  \\
OsiriX Lite \citep{PixmeoSARL2019} & 2004 & 2019-11 &  & 9 & 544304 &  & \checkmark &  &  \\
3D Slicer \citep{Kikinis2014} & 1998 & 2020-08 & \checkmark & 100 & 501451 & \checkmark & \checkmark & \checkmark &  \\
Drishti \citep{Limaye2012} & 2012 & 2020-08 &  & 1 & 268168 & \checkmark & \checkmark & \checkmark &  \\
Ginkgo CADx \citep{Wollny2020} & 2010 & 2019-05 &  & 3 & 257144 & \checkmark & \checkmark & \checkmark &  \\
GATE \citep{Jan2004} & 2011 & 2020-10 &  & 45 & 207122 &  & \checkmark & \checkmark &  \\
3DimViewer \citep{TESCAN2020} & ? & 2020-03 & \checkmark & 3 & 178065 & \checkmark & \checkmark &  &  \\
medInria \citep{Fillard2012} & 2009 & 2020-11 &  & 21 & 148924 & \checkmark & \checkmark & \checkmark &  \\
BioImage Suite Web \citep{Papademetris2005} & 2018 & 2020-10 & \checkmark & 13 & 139699 &
\checkmark & \checkmark & \checkmark & \checkmark \\
Weasis \citep{Roduit2021} & 2010 & 2020-08 &  & 8 & 123272 & \checkmark & \checkmark & \checkmark &  \\
AMIDE \citep{Loening2017} & 2006 & 2017-01 &  & 4 & 102827 & \checkmark & \checkmark & \checkmark &  \\
XMedCon \citep{Nolf2003} & 2000 & 2020-08 &  & 2 & 96767 & \checkmark & \checkmark & \checkmark &  \\
ITK-SNAP \citep{Yushkevich2006} & 2006 & 2020-06 & \checkmark & 13 & 88530 & \checkmark & \checkmark & \checkmark &  \\
Papaya \citep{UTHSCSA2019} & 2012 & 2019-05 &  & 9 & 71831 & \checkmark & \checkmark & \checkmark &  \\
OHIF Viewer \citep{Ziegler2020} & 2015 & 2020-10 &  & 76 & 63951 & \checkmark & \checkmark & \checkmark & \checkmark \\
SMILI \citep{Chandra2018} & 2014 & 2020-06 &  & 9 & 62626 & \checkmark & \checkmark & \checkmark &  \\
INVESALIUS 3 \citep{Amorim2015} & 2009 & 2020-09 &  & 10 & 48605 & \checkmark & \checkmark & \checkmark &  \\
dwv \citep{Martelli2021} & 2012 & 2020-09 &  & 22 & 47815 & \checkmark & \checkmark & \checkmark & \checkmark \\
DICOM Viewer \citep{Afsar2021} & 2018 & 2020-04 & \checkmark & 5 & 30761 & \checkmark & \checkmark & \checkmark &  \\
MicroView \citep{ParallaxInnovations2020} & 2015 & 2020-08 &  & 2 & 27470 & \checkmark & \checkmark & \checkmark &  \\
MatrixUser \citep{Liu2016} & 2013 & 2018-07 &  & 1 & 23121 & \checkmark & \checkmark & \checkmark &  \\
Slice:Drop \citep{Haehn2013} & 2012 & 2020-04 &  & 3 & 19020 & \checkmark & \checkmark & \checkmark & \checkmark \\
dicompyler \citep{Panchal2010} & 2009 & 2020-01 &  & 2 & 15941 & \checkmark & \checkmark &  &  \\
Fiji \citep{Schindelin2012} & 2011 & 2020-08 & \checkmark & 55 & 10833 & \checkmark & \checkmark & \checkmark &  \\
ImageJ \citep{Rueden2017} & 1997 & 2020-08 & \checkmark & 18 & 9681 & \checkmark & \checkmark & \checkmark &  \\
MRIcroGL \citep{Rorden2021} & 2015 & 2020-08 &  & 2 & 8493 & \checkmark & \checkmark & \checkmark &  \\
DicomBrowser \citep{Archie2012} & 2012 & 2020-08 &  & 3 & 5505 & \checkmark & \checkmark & \checkmark &  \\ \bottomrule
\end{tabular}
\caption{Final software list (sorted in descending order of the number of Lines
Of Code (LOC))}
\label{tab_final_list}
\end{table}

Figure \ref{fig_language} shows the primary languages versus the number of
projects using them.  The primary language is the language used for the majority
of the project's code; in most cases projects also use other languages.  The
most popular language is \CC, with almost 40\% of projects (11 of 29).  The two
least popular choices are Pascal and Matlab, with around 3\% of projects each
(1 of 29).

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.5]{figures/PrimaryLanguages.pdf}
\centering
\caption{\label{fig_language}Primary languages versus number of projects}
\end{figure}

\subsection{Installability} \label{sec_result_installability}

Figure \ref{fg_installability_scores} lists the installability scores.  We found
installation instructions for 16 projects. Among the ones without instructions,
\textit{BioImage Suite Web} and \textit{Slice:Drop} do not need installation,
since they are web applications. Installing 10 of the projects required extra
dependencies. Five of them are web applications (as shown in
Table~\ref{tab_final_list}) and depend on a browser; \textit{dwv}, \textit{OHIF
Viewer}, and \textit{GATE} needs extra dependencies to build; \textit{ImageJ}
and	\textit{Fiji} need an unzip tool; \textit{MatrixUser} is based on Matlab;
\textit{DICOM Viewer} needs to work on a Nextcloud platform.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/installability_scores.pdf}
\caption{AHP installability scores}
\label{fg_installability_scores}
\end{figure}

\textit{3D Slicer} has the highest score because it had easy to follow
installation instructions, and an automated, fast, frustration-free installation
process. The installer added all dependencies automatically and no errors
occurred during the installation and uninstallation steps. Many other software
packages also had installation instructions and automated installers.  We had no
trouble installing the following packages: \textit{INVESALIUS 3},
\textit{Gwyddion}, \textit{XMedCon}, and \textit{MicroView}. We determined their
scores based on the understandability of the instructions, installation steps,
and user experience. Since \textit{BioImage Suite Web} and \textit{Slice:Drop}
needed no installation, we gave them high scores. \textit{BioImage Suite Web}
also provided an option to download cache for offline usage, which was easy to
apply.

\textit{GATE}, \textit{dwv}, and \textit{DICOM Viewer} showed severe
installation problems. We were not able to install them, even after a reasonable
amount of time (2 hours).  For \textit{dwv} and \textit{GATE} we failed to build
from the source code, but we were able to proceed with measuring other qualities
using a deployed on-line version for \textit{dwv}, and a VM version for
\textit{GATE}. For \textit{DICOM Viewer} we could not install the NextCloud
dependency, and we did not have another option for running the software.
Therefore, for \textit{DICOM Viewer} we could not measure reliability or
robustness.  The other seven qualities could be measured, since they do not
require installation.

\textit{MatrixUser} has a lower score because it depends on Matlab. We assessed
the score from the point of view of a user that would have to install Matlab and
acquire a license.  Of course, for users that already work within Matlab, the
installability score should be higher.

\subsection{Correctness \& Verifiability} \label{sec_result_correctness_verifiability}

Figure~\ref{fg_correctness_verifiability_scores} shows the scores of correctness
and verifiability. Generally speaking, the packages with higher scores adopted
more techniques to improve correctness, and had better documentation for us to
verify against.  For instance, we looked for evidence of unit testing, since it
benefits most parts of the software's life cycle, such as designing, coding,
debugging, and optimization \citep{Hamill2004}.  We only found evidence of unit
testing for about half of the projects. We identified five projects using CI/CD
tools: \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, \textit{dwv}, and
\textit{OHIF Viewer}.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/correctness_verifiability_scores.pdf}
\caption{AHP correctness \& verifiability scores}
\label{fg_correctness_verifiability_scores}
\end{figure}

Even for some projects with well-organized documentation, requirements
specifications and theory manuals were still missing.  We could not identify
theory manuals for all projects, and we did not find requirements specifications
for most projects. The only requirements-related document we found was a road
map of \textit{3D Slicer}, which contained design requirements for upcoming
changes.

\subsection{Surface Reliability} \label{sec_result_reliability}

Figure~\ref{fg_reliability_scores} shows the AHP results.  As shown in
Section~\ref{sec_result_installability}, most of the software products did not
``break'' during installation, or did not need installation; \textit{dwv} and
\textit{GATE} broke in the building stage, and the processes were not
recoverable; we could not install the dependency for \textit{DICOM Viewer}. Of
the seven software packages with a getting started tutorial and operation steps
in the tutorial, most showed no error when we followed the steps. However,
\textit{GATE} could not open macro files and became unresponsive several times,
without any descriptive error message. When assessing robustness
(Section~\ref{sec_result_robustness}), we found that \textit{Drishti} crashed
when loading damaged image files, without showing any descriptive error message.
We did not find any problems with the on-line version of
\textit{dwv}.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/reliability_scores.pdf}
\caption{AHP surface reliability scores}
\label{fg_reliability_scores}
\end{figure}

\subsection{Surface Robustness} \label{sec_result_robustness}

Figure \ref{fg_robustness_scores} presents the scores for surface robustness.
The packages with higher scores elegantly handled unexpected/unanticipated
inputs, typically showing a clear error message. We may have underestimated the
score of \textit{OHIF Viewer}, since we needed further customization to load
data.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/robustness_scores.pdf}
\caption{AHP surface robustness scores}
\label{fg_robustness_scores}
\end{figure}

Digital Imaging and Communications in Medicine (DICOM) ``defines the formats for
medical images that can be exchanged with the data and quality necessary for
clinical use'' \citep{MITA2021}. According to their documentation, all 29
software packages should support the DICOM standard. To test robustness, we
prepared two types of image files: correct and incorrect formats (with the
incorrect format created by relabelled a text file to have the ``.dcm''
extension).  All software packages loaded the correct format image, except for
\textit{GATE}, which failed for unknown reasons.  For the broken format,
\textit{MatrixUser}, \textit{dwv}, and \textit{Slice:Drop} ignored the incorrect
format of the file and loaded it regardless. They did not show any error message
and displayed a blank image. \textit{MRIcroGL} behaved similarly except that it
showed a meaningless image. \textit{Drishti} successfully detected the broken
format of the file, but the software crashed as a result.

\subsection{Surface Usability} \label{sec_result_usability}

Figure~\ref{fg_usability_scores} shows the AHP scores for surface usability. The
software with higher scores usually provided both comprehensive documented
guidance and a good user experience. \textit{INVESALIUS 3} provided an excellent
example of a detailed and precise user manual. \textit{GATE} also provided
numerous documents, but unfortunately we had difficulty understanding and using
them. We found getting started tutorials for only 11 projects, but a user manual
for 22 projects. \textit{MRIcroGL} was the only project that explicitly
documented expected user characteristics.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/usability_scores.pdf}
\caption{AHP surface usability scores}
\label{fg_usability_scores}
\end{figure}

\subsection{Maintainability} \label{sec_score_maintainability}

Figure~\ref{fg_maintainability_scores} shows the ranking results for
maintainability. We gave \textit{3D Slicer} the highest score because we found
it to have the most comprehensive artifacts. For example, as far as we could
find, only a few of the 29 projects had a product, developer's manual, or API
documentation, and only \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}
included all three documents. Moreover, \textit{3D Slicer} has a much higher
percentage of closed issues (91.65\%) compared to \textit{ImageJ} (52.49\%) and
\textit{Fiji} (63.79\%). Table~\ref{tab_maintainability_docs} shows which
projects had these documents, in descending order of their maintainability
scores.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/maintainability_scores.pdf}
\caption{AHP maintainability scores}
\label{fg_maintainability_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lccc}
\toprule
\multicolumn{1}{c}{Software} & Prod.\ Roadmap & Dev.\ Manual & API Doc. \\ 
\midrule
3D Slicer & \checkmark & \checkmark & \checkmark \\
ImageJ & \checkmark & \checkmark & \checkmark \\
Weasis &  & \checkmark &  \\
OHIF Viewer &  & \checkmark & \checkmark \\
Fiji & \checkmark & \checkmark & \checkmark \\
ParaView & \checkmark &  &  \\
SMILI &  &  & \checkmark \\
medInria &  & \checkmark &  \\
INVESALIUS 3 & \checkmark &  &  \\
dwv &  &  & \checkmark \\
BioImage Suite Web &  & \checkmark &  \\
Gwyddion &  & \checkmark & \checkmark \\ 
\bottomrule
\end{tabular}
\caption{Software with the maintainability documents (listed in descending order of 
maintainability score)}
\label{tab_maintainability_docs}
\end{table}

Twenty-seven of the 29 projects used git as the version control tool, with 24 of these
using GitHub. \textit{AMIDE} used Mercurial and \textit{Gwyddion} used
Subversion. \textit{XMedCon}, \textit{AMIDE}, and \textit{Gwyddion} used
SourceForge. \textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

\subsection{Reusability} \label{sec_result_reusability}

Figure~\ref{fg_reusability_scores} shows the AHP results for reusability. As
described in Section~\ref{sec_grading_software}, we gave higher scores to the
projects with API documentation. As shown in
Table~\ref{tab_maintainability_docs}, seven projects had API documents. We also
assumed that projects with more code files and less LOC per code file are more
reusable. Table \ref{tab_loc_per_file} shows the number of text-based files by
project, which we used to approximate the number of code files. The table also
lists the total number of lines (including comments and blanks), LOC, and
average LOC per file. We arranged the items in descending order of their
reusability scores.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/reusability_scores.pdf}
\caption{AHP reusability scores}
\label{fg_reusability_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
\multirow{2}{*}{Software} & \multirow{2}{*}{Text Files} & \multirow{2}{*}{Total Lines} & \multirow{2}{*}{LOC} & \multirow{2}{*}{LOC/file} \\
 &  &  &  &  \\ 
\midrule
OHIF Viewer & 1162 & 86306 & 63951 & 55 \\
3D Slicer & 3386 & 709143 & 501451 & 148 \\
Gwyddion & 2060 & 787966 & 643427 & 312 \\
ParaView & 5556 & 1276863 & 886326 & 160 \\
OsiriX Lite & 2270 & 873025 & 544304 & 240 \\
Horos & 2346 & 912496 & 561617 & 239 \\
medInria & 1678 & 214607 & 148924 & 89 \\
Weasis & 1027 & 156551 & 123272 & 120 \\
BioImage Suite Web & 931 & 203810 & 139699 & 150 \\
GATE & 1720 & 311703 & 207122 & 120 \\
Ginkgo CADx & 974 & 361207 & 257144 & 264 \\
SMILI & 275 & 90146 & 62626 & 228 \\
Fiji & 136 & 13764 & 10833 & 80 \\
Drishti & 757 & 345225 & 268168 & 354 \\
ITK-SNAP & 677 & 139880 & 88530 & 131 \\
3DimViewer & 730 & 240627 & 178065 & 244 \\
DICOM Viewer & 302 & 34701 & 30761 & 102 \\
ImageJ & 40 & 10740 & 9681 & 242 \\
dwv & 188 & 71099 & 47815 & 254 \\
MatrixUser & 216 & 31336 & 23121 & 107 \\
INVESALIUS 3 & 156 & 59328 & 48605 & 312 \\
AMIDE & 183 & 139658 & 102827 & 562 \\
Papaya & 110 & 95594 & 71831 & 653 \\
MicroView & 137 & 36173 & 27470 & 201 \\
XMedCon & 202 & 129991 & 96767 & 479 \\
MRIcroGL & 97 & 50445 & 8493 & 88 \\
Slice:Drop & 77 & 25720 & 19020 & 247 \\
DicomBrowser & 54 & 7375 & 5505 & 102 \\
dicompyler & 48 & 19201 & 15941 & 332 \\ 
\bottomrule
\end{tabular}
\caption{Number of files and lines (sorted in descending order of reusability
scores)}
\label{tab_loc_per_file}
\end{table}

\subsection{Surface Understandability} \label{sec_result_understandability}

Figure~\ref{fg_surface_understandability_scores} shows the scores for surface
understandability. All projects had a consistent coding style with parameters in
the same order for all functions; modularized code; and, clear comments, indicating
what is done, not how. However, we only found explicit identification of a
coding standard for 3 out of the 29: \textit{3D Slicer}, \textit{Weasis}, and
\textit{ImageJ}. We also found hard-coded constants (rather than symbolic
constants) in \textit{medInria}, \textit{dicompyler}, \textit{MicroView}, and
\textit{Papaya}. We did not find any reference to the algorithms used in
projects \textit{XMedCon}, \textit{DicomBrowser}, \textit{3DimViewer},
\textit{BioImage Suite Web}, \textit{Slice:Drop}, \textit{MatrixUser},
\textit{DICOM Viewer}, \textit{dicompyler}, and \textit{Papaya}.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/understandability_scores.pdf}
\caption{AHP surface understandability scores}
\label{fg_surface_understandability_scores}
\end{figure}

\subsection{Visibility/Transparency} \label{sec_result_visibility_transparency}

Figure~\ref{fg_visibility_transparency_scores} shows the AHP scores for
visibility/transparency. Generally speaking, the teams that actively documented
their development process and plans scored higher.
Table~\ref{tab_Visibility/Transparency_docs} shows the projects that had
documents for the development process, project status, development environment,
and release notes, in descending order of their visibility/transparency
scores.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/visibility_transparency_scores.pdf}
\caption{AHP visibility/transparency scores}
\label{fg_visibility_transparency_scores}
\end{figure}

\begin{table}[!ht]
\centering
\begin{tabular}{lllll}
\toprule
Software & Dev.\ Process & Proj.\ Status & Dev.\ Env. & Rls.\ Notes \\ 
\midrule
3D Slicer & \checkmark & \checkmark & \checkmark & \checkmark \\
ImageJ & \checkmark & \checkmark & \checkmark & \checkmark \\
Fiji & \checkmark & \checkmark & \checkmark &  \\
MRIcroGL &  &  &  & \checkmark \\
Weasis &  &  & \checkmark & \checkmark \\
ParaView &  & \checkmark &  &  \\
OHIF Viewer &  &  & \checkmark & \checkmark \\
DICOM Viewer &  &  & \checkmark & \checkmark \\
medInria &  &  & \checkmark & \checkmark \\
SMILI &  &  &  & \checkmark \\
Drishti &  &  &  & \checkmark \\
INVESALIUS 3 &  &  &  & \checkmark \\
OsiriX Lite &  &  &  & \checkmark \\
GATE &  &  &  & \checkmark \\
MicroView &  &  &  & \checkmark \\
MatrixUser &  &  &  & \checkmark \\
BioImage Suite Web &  &  & \checkmark &  \\
ITK-SNAP &  &  &  & \checkmark \\
Horos &  &  &  & \checkmark \\
dwv &  &  &  & \checkmark \\
Gwyddion &  &  &  & \checkmark \\ 
\bottomrule
\end{tabular}
\caption{Software with visibility/transparency related documents (listed in
descending order of visibility/transparency score)}
\label{tab_Visibility/Transparency_docs}
\end{table}

\subsection{Overall Scores} \label{Sec_OverallQ}

As described in Section~\ref{sec_AHP}, for our AHP measurements, we have nine
criteria (qualities) and 29 alternatives (software packages). In the absence of
a specific real world context, we assumed all nine qualities are equally
important. Figure~\ref{fg_overall_scores} shows the overall scores in descending
order. Since we produced the scores from the AHP process, the total sum of the
29 scores is precisely 1.0.

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

\label{fg_overall_scores}
\end{figure}

The top four software products \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, and \textit{OHIF Viewer} have higher scores in most criteria.
\textit{3D Slicer} has a score in the top two for all qualities; \textit{ImageJ}
ranks near the top for all qualities, except for correctness \& verifiability.
\textit{OHIF Viewer} and \textit{Fiji} have similar overall scores, with
\textit{Fiji} doing better in installability and \textit{OHIF Viewer} doing
better in correctness \& verifiability.  Given the installation problems, we may
have underestimated the scores on reliability and robustness for \textit{DICOM
Viewer}, but we compared it equally for the other seven qualities.

\section{Comparison to Community Ranking} \label{Sec_VsCommunityRanking}

To address~\rqref{RQ_CompareHQ2Popular} about how our ranking compares to the
popularity of projects as judged by the scientific community, we make two
comparisons:
\begin{itemize}
\item A comparison of our ranking (from Section~\ref{ch_results}) with the
community ratings on GitHub, as shown by GitHub stars, number of forks, and
number of people watching the projects; and,
\item A comparison of top-rated software from our methodology with the top
recommendations from our domain experts (as mentioned in
Section~\ref{sec_software_selection}).
\end{itemize}

Table~\ref{tab_ranking_vs_GitHub} shows our ranking of the 29 MI projects, and
their GitHub metrics, if applicable. As mentioned in
Section~\ref{sec_score_maintainability}, 24 projects used GitHub. Since GitHub
repositories have different creation dates, we collect the number of months each
stayed on GitHub, and calculate the average number of new stars, people
watching, and forks per 12 months. Section~\ref{sec_grading_software} describes
the method of getting the creation date.  The items in
Table~\ref{tab_ranking_vs_GitHub} are listed in descending order of the average
number of new stars per year.  The non-GitHub items are listed in the order of
our ranking.  We collected all GitHub statistics in July 2021.

Generally speaking, most of the top-ranking MI software projects also received
greater attention and popularity on GitHub. Between our ranking and the GitHub
stars-per-year ranking, four of the top five software projects appear in both
lists. Our top five packages are scattered among the first eight positions on the
GitHub list. However, as discussed below there are discrepancies between the two
lists.

In some cases projects are popular in the community, but were assigned a low
rank by our methodology.  This is the case for \textit{dwv}. The reason for the
low ranking is that, as mentioned in Section~\ref{sec_result_installability}, we
failed to build it locally, and used the test version on its websites for the
measurements. We followed the instructions and tried to run the command ``yarn
run test'' locally, which did not work. In addition, the test version did not
detect a broken DICOM file and displayed a blank image as described in
Section~\ref{sec_result_robustness}. We might underestimate the scores for
\textit{dwv} due to uncommon technical issues.

We also ranked \textit{DICOM Viewer} much lower than its popularity. As
mentioned in Section~\ref{sec_result_installability}, it depended on the
NextCloud platform that we could not successfully install. Thus, we might
underestimate the scores of its surface reliability and surface robustness.

Further reason for discrepancies between our ranking and the community's ranking
is that we weighted all qualities equally. This is not likely how users
implicitly rank the different qualities. As a result, some projects with high
community popularity may have scored lower with our method because of a
relatively higher (compared to the scientific community's implicit ranking)
weighting of the poor scores for some qualities. A further explanation for
discrepancies between our measures and the star measures may also be due to
inaccuracy with using stars to approximate popularity.  Stars are not an ideal
measure because stars represent the community's feeling in the past more than
they measure current preferences \citep{Szulik2017}.  The issue with stars is
that they tend only to be added, not removed.  A final reason for
inconsistencies between our ranking and the community's ranking is that, as for
consumer products, more factors influence popularity than just quality.

>>>>

\documentclass[11pt]{article}

\usepackage{authblk}
\usepackage{fancyhdr}
\usepackage{float}

\usepackage[round, sort]{natbib}

\usepackage{amssymb}
\usepackage{amsmath,amsthm}

@@

\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{pbox}

\oddsidemargin 0mm
\evensidemargin 0mm
\textwidth 165mm
\textheight 205mm

\pagestyle{fancy}
\lhead{\leftmark}
\chead{}
\rhead{\rightmark}
\cfoot{\thepage}

@@

@@

\title{State of the Practice for Medical Imaging Software Based on Open Source Repositories}

\author[1,*]{W.\ Spencer Smith}
\author[1]{Ao Dong}
\author[1]{Jacques Carette}
\author[2]{Michael D.\ Noseworthy}

\affil[1]{McMaster University, Computing and Software Department, {Canada}}
\affil[2]{McMaster University, Electrical \& Computer Engineering Department,
{Canada}}
\affil[*]{Corresponding Author}

\maketitle

% RQ5 to RQ10
% Some of Section 1 (Introduction)
% Sections 7 to 10 (includes use of tools, some of which was based on interviews with developers)
% Some of Sections 11 to 13 (Threats to Validity, Future Work, Conclusions)

@@

\noindent \emph{Keywords:}
	medical imaging, research software, software engineering, software
	quality, analytic hierarchy process, developer interviews

@@

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[scale=0.25]{MPR.png}        
    \end{center}
    \caption{Example brain image showing a multi-planar reformat using Horos
	(free open-source medical imaging/DICOM viewer for OSX, based on OsiriX)}
    \label{Fig_Example}
\end{figure}

@@

We designed a general methodology to assess the state of the practice for
research software \citep{SmithEtAl2021, SmithAndMichalski2022}. Details can be
found in Section~\ref{ch_methods}.  Our methodology has been applied to MI
software \citep{Dong2021} and Lattice Boltzmann Solvers \citep{Michalski2021,
SmithEtAl2024}.  This methodology builds off prior work to assess the state of
the practice for such domains as Geographic Information Systems
\citep{smith2018state}, Mesh Generators \citep{smith2016state}, Seismology
software \citep{Smith2018Seismology}, and Statistical software for psychology
\citep{smith2018statistical}.  In keeping with the previous methodology, we have
maintained the constraint that the work load for measuring a given domain should
be feasible for a team as small as one person, and for a short time, ideally
around a person month of effort. We consider a person month as $20$ working days
($4$ weeks in a month, with $5$ days of work per week) at $8$ person hours per
day, or $20 \times 8 = 160$ person hours.

@@

\begin{figure}[!ht]
\includegraphics[scale=0.66]{template.pdf}
\caption{Grading template example}
\label{fg_grading_template_example}
\end{figure}

@@
